### 2022-03-08 paper/images/abstract.tex
let's hope I didn't just dox myself

% this is SO DUMB
\documentclass{article}
\usepackage{multicol}

\begin{document}
\begin{multicols}{3}
    \centering
    \textbf{Purpose} \\
    \raggedright To explore quantum computing and development of NISQ devices by implementing the theoretical ``quantum bogosort'' gag algorithm. \\
    \columnbreak
    \centering
    \textbf{Procedure} \\
    \raggedright Implemented with Qiskit and tested on Aer simulator and IBMQ Lima (5 qubit machine). \\
    \columnbreak
    \centering
    \textbf{Conclusion} \\
    \raggedright Quantum bogosort is a funky way to introduce quantum computing fundamentals to beginners. Novel QRNG algorithm provides insight into crafting arbitrary qubit states but is not practically useful.
\end{multicols}
\end{document}